---
title: "how-to-reproject-to-a-custom-plate-carree"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-reproject-to-a-custom-plate-carree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(geosparklines)
library(readr)
library(magrittr)
library(dplyr)
library(sf)
```

{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prepare your Data
First, download the world COVID-19 cases from John Hopkins Institute :
```{r}
f <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
```

Then, let's select the relevant informations :
```{r}
coords <- f[, c("Long", "Lat")]
col_dates <- grep("[0-9]*/[0-9]*/[0-9]*", names(f)) # select all column dates
my_stats <- f[, col_dates]
```

## Reproject to Plate Carrée
WGS-84 might not be the appropriate Coordinate Reference System as it distorts the countries and the distances on a world map. For instance, countries seem bigger than they are as we move upwards or downwards.

This projection might be prejudiciable to the reading of your sparklines.

That's why an Equal-Distance projection may be more appropriate. For example, Plate Carrée projection preserves proportion of countries.

### A custom Plate Carrée
Nevertheless, you might want to change the CRS bounds of your projection, because if you have big lines, high and wide, they might overflow with your map.

I created my own Plate Carrée CRS with shifted bounds (see lon_1, lat_1, lon_2 and lat_2 paramaters) :
```{r}
library(sp)

xmin <- -1000
ymin <- -1000
xmax <- 1000
ymax <- 1000

crs_string <- sprintf("+proj=eqc +lat_ts=0 +lat_1=%d +lon_1=%d  +lat_2=%d +lon_2=%d  +lat_0=0 +lon_0=0  +x_0=0 +y_0=0  +datum=WGS84 +units=m +no_defs", ymin, xmin, ymax, xmax )

my_crs <- crs_string %>% CRS
```

Let's reproject data
```{r message = FALSE}
coords_pc <- lapply(1:nrow(coords), function(x) {
  st_point(as.numeric(coords[x, ])) %>% 
  st_sfc %>% 
  st_set_crs(4326) %>% st_transform(my_crs) %>% 
  st_sfc %>% 
  st_coordinates
}) %>% do.call(rbind, .)

```

Now, you can use it the same way as before for your data. But you'll need to adjust width and height according to this new coordinate reference system
```{r}
geosparks <- geospark(coords_pc, my_stats, width=1000000, height=10000000, mode = "log")
plot(geosparks)
st_geometry(f) <- geosparks
```
