---
title: "How to use geosparklines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comment-utiliser-mon-package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(geosparklines)
library(readr)
library(magrittr)
library(dplyr)
library(sf)
library(sp)
```

## Prepare your Data
First, download the world COVID-19 cases from John Hopkins Institute
```{r}
f <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
```

You'll need :

- some coordinates for the source points
- the stat values (here in wide column-wise format, not long row-wise format) which will draw the lines
```{r}
coords <- f[, c("Long", "Lat")]
my_stats <- f[, 5:ncol(f)]
```

## Create your GeoSparkLines
Define the width and height in spatial units, for your lines and the mode : `normal` or `log` for your values

```{r}
## Whole world
geosparks <- geospark(coords, my_stats, width=20, height=20, mode = "log")
plot(st_geometry(geosparks))
```

You can add your GeoSparkLines to your data :
```{r}
st_geometry(f) <- geosparks
```

Let's plot it, based on a log value of the last day cases

```{r message=FALSE}
## Plot
plot(f %>%
       mutate(v = ifelse(.$`4/17/20` <= 0, 0, log(.$`4/17/20`))) %>%
       select(v))
```

### Export your work
...for further styling, for example in QGIS
```{r eval=FALSE}
st_write(f, "data/world_sparklines.gpkg", delete_dsn = TRUE)

```


## Reproject to Plate Carrée
Notice the Coordinate Reference System is not appropriate for our purpose because it's WGS84 and it distorts the lines to the extreme as we go North or South.

So, let's the adopt a Plate Carrée projection. I created my own CRS with shifted bounds (see lon_1, lat_1, lon_2 and lat_2 paramaters) :
```{r}
library(sp)
crs_string <- "+proj=eqc +lat_ts=0 +lat_1=-1000 +lon_1=-1000  +lat_2=1000 +lon_2=1000  +lat_0=0 +lon_0=0  +x_0=0 +y_0=0  +datum=WGS84 +units=m +no_defs"

my_crs <- crs_string %>% CRS
```

Let's reproject data
```{r}
coords_pc <- lapply(1:nrow(coords), function(x) st_point(as.numeric(coords[x, ]))) %>% st_sfc %>% st_set_crs(4326) %>% st_transform(my_crs) %>% st_sfc %>% st_coordinates

```

Now, you can use it the same way as before for your data. But you'll need to adjust width and height according to the coordinate reference system
```{r}
geosparks <- geospark(coords_pc, my_stats, width=1000000, height=1000000, mode = "log")
plot(geosparks)
```

## Individual sparklines
You may want to create sparklines individually, for example to pilot the individual heights and widths of your sparklines based on statistical values.

You can do it with `one_geospark`.

Here is an example on France.

```{r}
i <- which(f$`Country/Region`=="France" & is.na(f$`Province/State`))
v <- my_stats[i, ] %>% as.numeric
my_coords <- coords[i, ] %>% as.matrix
sp_line <- one_geospark(my_coords, v, width=20, height=20, mode="log")
plot(st_geometry(sp_line))
```

By default, the maximum value which defines the height will be the maximum value of the values you give to the function.

You can put a custom max value. Here, I use the maximum value of the global dataset
```{r}
sp_line <- one_geospark(my_coords, v, max_v = max(my_stats), width=20, height=20, mode="log")
plot(st_geometry(sp_line))
```
