---
title: "How to use geosparklines"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{comment-utiliser-mon-package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(geosparklines)
library(readr)
library(magrittr)
library(dplyr)
library(sf)
```

## Prepare your Data
You'll need :

- some coordinates for the source points
- the stat values (here in wide column-wise format, not long row-wise format) which will draw the lines

First, download the world COVID-19 cases from John Hopkins Institute :
```{r message = FALSE}
f <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
```

Then, let's select the relevant informations :
```{r}
coords <- f[, c("Long", "Lat")]
col_dates <- grep("[0-9]*/[0-9]*/[0-9]*", names(f)) # select all column dates
my_stats <- data.frame(f)[, col_dates]
```

## Create your GeoSparkLines
Define the width and height in spatial units of your lines and the transformation mode : `normal` or `log` for your values

```{r}
## Whole world
geosparks <- geospark(coords, my_stats, width=4, height=2, direction = "ne", mode = "log") %>% st_set_crs(4326)
print(class(geosparks))
```

You can add your GeoSparkLines to your data :
```{r}
st_geometry(f) <- geosparks
```

Let's do some simple plotting :
```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(rgeos)
world <- ne_countries(scale = "medium", returnclass = "sf")

last_day <- names(f)[col_dates %>% tail(1)]

ggplot() +
    geom_sf(data=world) +
    geom_sf(data=f, col="red") +
    coord_sf(xlim = c(-20, 30), ylim = c(25, 60), expand = FALSE) +
    ggtitle(sprintf("Covid-19 cases sparklines (log-transformed)\n%s", last_day))
```

### Export your work
...for further styling, for example in QGIS
```{r eval=FALSE}
st_write(f, "world_sparklines.gpkg", delete_dsn = TRUE)

```

## Individual sparklines
You might want to create sparklines individually, for example to pilot the individual heights and widths of your sparklines more precisely

You can do it with `singleGeospark`.

Here is an example on France.

```{r}
i <- which(f$`Country/Region`=="France" & is.na(f$`Province/State`))
v <- my_stats[i, ] %>% as.numeric
my_coords <- coords[i, ] %>% as.matrix
sp_line <- singleGeospark(my_coords, v, width=20, height=20, mode="log")
plot(st_geometry(sp_line))
```

By default, the maximum value which defines the height will be the maximum value of the values you give to the function.

You can put a custom max value. Here, I use the maximum value of the global dataset
```{r}
sp_line <- singleGeospark(my_coords, v, max_v = max(my_stats), width=20, height=20, mode="log")
plot(st_geometry(sp_line))
```
